---
title: "Permutation Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(ggmosaic)
knitr::opts_chunk$set(echo = FALSE)

library(palmerpenguins)
my_penguins <- penguins %>% 
  filter(species == "Adelie",
         !is.na(sex), 
         island == "Torgersen") 
library(infer)
```

# Permutation quiz

## Background


Some species display sexual size dimorphism -- in which one sex is on average larger than the other. Such a pattern can tell us about the species' ecology and mating habits. Do penguins display this sex difference in size? To start to answer thus, let's look at a subset of the palmerpenguins data set, which I'll call `my_penguins`.


```{r, eval = FALSE, echo=TRUE}
library(palmerpenguins)
my_penguins <- penguins %>% 
  filter(species == "Adelie",
         !is.na(sex), 
         island == "Torgersen")  
``` 

#### `glimpse()`

First, take a `glimpse()` at the `my_penguins`  dataset

```{r glimpse, exercise=TRUE, eval = FALSE}
____(___)
```

```{r glimpse-solution, eval = FALSE}
glimpse(my_penguins)
```

#### Visualize

First let's visualize body size by sex

```{r plot, exercise=TRUE, eval = FALSE}
ggplot(data = my_penguins, aes(x = __, y = ___, color = sex))+
  ____()  +
  stat_summary(fun.data = "mean_Cl_boot", color = "black")
```

```{r plot-solution}
ggplot(data = my_penguins, aes(x = sex, y = body_mass_g, color = sex))+
  geom_jitter(show.legend = FALSE)  +
  stat_summary(fun.data = "mean_cl_boot", color = "black")
```


Now lets estimate the sex difference in `body_mass_g` (male mass - female) from the data, and the standard deviation within each sex.



```{r est, exercise=TRUE, eval = FALSE}
# diff in means
my_penguins %>%
  group_by(  ____ ) %>%
  summarize(mean_body = ___(___)) %>%
  pivot_wider() %>%
  pivot_wider(names_from = sex, values_from = mean_body) %>%
  mutate(sex_diff = __ - __)

# sd of each
my_penguins %>%
  group_by(  ____ ) %>%
  summarize(sd_body = ___(___)) 
```

```{r est-solution}
# diff in means
my_penguins %>%
  group_by(  sex ) %>%
  summarize(mean_body = mean(body_mass_g)) %>%
  pivot_wider(names_from = sex, values_from = mean_body) %>%
  mutate(sex_diff = male - female)

# sd of each
my_penguins %>%
  group_by( sex ) %>%
  summarize(sd_body = sd(body_mass_g)) 
```

```{r obs, echo = FALSE}
question_text("The estimated difference between male and female bedy weight of Adelie penguins  from Torgersen is (round to nearest integer)",
 answer("639", correct = TRUE),
 allow_retry = TRUE,   
  correct = random_praise(),
  incorrect = random_encouragement()
)
```

#### Now lets think about bootrapping and permuting   


```{r  bootrappingAndPermuting, echo=FALSE}
quiz(caption = " bootrapping and permuting",
    question("Every bootstrap replicate will have  (select all that apply)",
    answer("Identical values for mean male mass"),
    answer("Identical values for mean female mass"),
    answer("A variable sex difference in mean body mass, deviating from the sample mean by sampling error", correct = TRUE),     
    answer("A variable sex difference in mean body mass, deviating from the proposed mean undr the null by sampling error"),  
    allow_retry = TRUE,
    incorrect = "Try again!"
  ),
        question("Every permuted replicate will have (select all that apply)",
    answer("Identical values for mean male mass", correct = TRUE),
    answer("Identical values for mean female mass", correct = TRUE),
    answer("A variable sex difference in mean body mass, deviating from the sample mean by sampling error"),     
    answer("A variable sex difference in mean body mass, deviating from the proposed mean undr the null by sampling error", correct = TRUE),  
    allow_retry = TRUE,
    incorrect = "Try again!"
  )
)
```


#### Bootstrap to  Estimate Uncertainty 

Say we want to estimate uncertainty in the sex difference in `body_mass_g` (male mass - female) from the data. Adapt the frog code,  below, to estimate uncertainty in the sex difference in body mass in `my_penguins` as the standard error and 95% confidence interval.      



```{r bootpenguins, exercise=TRUE, eval = FALSE}
n_reps <- 1000
# number nonpreferred

penguin_boot_sexdiff <- replicate(n = n_reps, simplify = FALSE, 
  expr = my_penguins %>%
    dplyr::select(sex, body_mass_g) %>%
    group_by(sex) %>%
    ___(___ = __, replace = ___) %>% ## resample with replacement to bootstrap
    group_by(sex) %>%
    summarize(mean_body = mean(body_mass_g)) %>%
    pivot_wider(names_from = sex, values_from = mean_body) %>%
    mutate(sex_diff = male - female))  %>%
  bind_rows()

penguin_boot_sexdiff  %>%
  summarise(se       = __(__),
            lower_CI = __(__, prob = __),
            upper_CI = __(__, prob = __))
```


```{r bootq, echo=FALSE}
question("What plausible parameter values for the sex difference in body size (male - female) are contained in the 95% confidence interval? (check all that apply)",
  answer("300"),
  answer("400"),
  answer("500", correct = TRUE),
  answer("600", correct = TRUE),
  answer("700", correct = TRUE),
  allow_retry = TRUE,
    correct = random_praise(),
  incorrect = random_encouragement()
)
```


#### Hypotheses    


```{r hypoth, echo=FALSE}
quiz(caption = "Hypotheses",
    question("What is the null hypothesis?",
    answer("Mean male and female body mass of Adelie penguins from Torgersen do not differ", correct = TRUE),
     answer("Mean body mass of male Adelie penguins from Torgersen is greater   than mean female body mass"),
      answer("Mean body mass of male Adelie penguins from Torgersen is smaller than mean female body mass"),     
    answer("Mean male and female body mass of Adelie penguins from Torgersen  differs"),
    answer("Mean body mass of male Adelie penguins from Torgersen is 638.9493
 grams  greater  than mean female body mass"),
    allow_retry = TRUE,
    incorrect = "Try again!"
  ),
    question("What is the alternative hypothesis?",
    answer("Mean male and female body mass of Adelie penguins from Torgersen do not differ"),
     answer("Mean body mass of male Adelie penguins from Torgersen is greater   than mean female body mass"),
      answer("Mean body mass of male Adelie penguins from Torgersen is smaller than mean female body mass"),     
    answer("Mean male and female body mass of Adelie penguins from Torgersen  differs", correct = TRUE),
    answer("Mean body mass of male Adelie penguins from Torgersen is 638.9493
 gram's  greater  than mean female body mass"),
    allow_retry = TRUE,
    incorrect = "Try again!"
  )
)
```


#### Permute to  Quantify Surprise

Say we wanted to  to estimate uncertainty in the sex difference in `body_mass_g` (male mass - female) from the data. Permute sex to test this null and estimate a p-value.    

```{r perm, exercise=TRUE, eval = FALSE}
n_reps <- 1000

penguin_perm_sexdiff <- replicate(n = n_reps, simplify = FALSE, 
  expr = my_penguins %>%
   dplyr::select(sex, body_mass_g) %>%
   mutate(perm_sex = sample(__, replace = __)) %>% ## resample with replacement to bootstrap
     group_by(__) %>%
     summarize(mean_body = mean(body_mass_g)) %>%
     pivot_wider(names_from = __, values_from = mean_body) %>%
     mutate(sex_diff_mass = male - female))%>% 
  bind_rows()

p.val <- penguin_perm_sexdiff  %>%
  mutate(as_or_more_extreme =  __(___) __ abs(__)) %>%
  summarise(p_val = mean(as_or_more_extreme))


p.val
```

```{r perm-solution}
n_reps <- 1000

penguin_perm_sexdiff <- replicate(n = n_reps, simplify = FALSE, 
  expr = my_penguins %>%
   dplyr::select(sex, body_mass_g) %>%
   mutate(perm_sex = sample(sex, replace = FALSE)) %>% ## resample with replacement to bootstrap
     group_by(perm_sex) %>%
     summarize(mean_body = mean(body_mass_g)) %>%
     pivot_wider(names_from = perm_sex, values_from = mean_body) %>%
     mutate(sex_diff_mass = male - female))%>% 
  bind_rows()


p.val <- penguin_perm_sexdiff %>%
  mutate(as_or_more_extreme =  abs(sex_diff_mass) >= abs(638.9493)) %>%
  summarise(p_val = mean(as_or_more_extreme))


p.val
```


```{r pvalquiz, echo=FALSE}
quiz(caption = "Interpetting p-values",
    question("What is our P value (approximately)?",
    answer("nearly zero", correct = TRUE),
    answer("Approx 0.01"),
    answer("Approx 0.1"),
    answer("Approx 0.5"),
    answer("nearly 1"),
    allow_retry = TRUE,
    incorrect = "Try again!"
  ),
  question("What do we do to the null hypothesis?",
    answer("Reject it", correct = TRUE),
    answer("Fail to reject it"),
    allow_retry = TRUE,
    incorrect = "Try again!"
  ),
    question("We can't **KNOW** if the null hypothesis is true or false, but which do you think is far more likely?)",
    answer("Null hypothesis is true"),
    answer("Null hypothesis is false",  correct = TRUE),
    allow_retry = TRUE,
    incorrect = "Try again!"
  )
)
```


**Be prepared to explain your answer to the question above**    
