---
title: "Probability simulations in R"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(ggmosaic)
knitr::opts_chunk$set(echo = FALSE)
```

# Probabilistic simulation quiz  

## Background

Imagine there is some disease spreading, and people concerned about the disease take a test with the following behavior:   

- The test correctly reports someone uninfected as uninfected in 0.995 of tests.  
- The test correctly reports someone infected as infected in 0.94 of tests.   

### Simulate sickness 

To better understand this case, lets simulate ten thousand tests in a population where **2%** of people getting tested are sick, and find the proportion of people that are sick. 


```{r sim1, exercise=TRUE, eval = FALSE}
n_tests  <- 10000
p_sick   <- 0.02

sickness <- tibble(
  actual_status = sample(c("sick","healthy"),
                         size = ___,
                         replace = TRUE,
                         prob = c(___, 1 - ___)))
)

sickness %>% 
  summarise(prop_sick = mean( actual_status == "___"))
```

```{r sim1-solution}
n_tests  <- 10000
p_sick   <- 0.02
sickness <- tibble(
  actual_status = sample(c("sick","healthy"),
                         size = n_tests,
                         replace = TRUE,
                         prob = c(p_sick, 1-p_sick))
)

sickness %>% 
  summarise(prop_sick = mean( actual_status == "sick"))
```

### Uncertainty

Hit the `run code` button a bunch of times to get a sense of the expected  variability across samples of size ten thousand. Then increase the number of tests to one million, and hit `run code` a few more times

```{r samplesize, echo=FALSE}
question("On average which sample size has a proportion of sick people closest to the true population parameter?",
  answer("Ten thousand"),
  answer("One million", correct = TRUE),
  answer("No discernable pattern"),
  allow_retry = TRUE
)
```

### Sampling distribution   

Generate a sampling distribution by running  ten thousand tests in one thousand replicate universes. Then plot the sampling distribution for the proportion of people who are sick and compute the standard error for the expected proportion of people (in a sample of ten thousand) who will be sick. 

```{r simreps, exercise=TRUE, eval = FALSE}
n_tests  <- 10000
p_sick   <- 0.02
n_reps   <- 1000

sickness_replicates <- tibble(
  actual_status = sample(c("sick","healthy"),
                         size = ___ * ___,
                         replace = TRUE,
                         prob = c(___, 1-___)),
  rep_id = rep(1:n_reps, each = ___))

sickness_sampling_dist <- sickness_replicates %>%
  group_by(___) %>% 
  ___(prop_sick = ___( ___ == ___))

# plot the sampling distribution
ggplot(data = sickness_sampling_dist, aes(x = prop_sick)) +
  geom_histogram(binwidth = 1e-3, color = "white", fill = "firebrick")

# find the standard error for the proportion of  people who are sick

```

```{r simreps-solution}
n_tests  <- 10000
p_sick   <- 0.02
n_reps   <- 1000

sickness_replicates <- tibble(
  actual_status = sample(c("sick","healthy"),
                         size = n_tests * n_reps,
                         replace = TRUE,
                         prob = c(p_sick, 1-p_sick)),
  rep_id = rep(1:n_reps, each = n_tests))

sickness_sampling_dist <- sickness_replicates %>%
  group_by(rep_id) %>% 
  summarise(prop_sick = mean( actual_status == "sick"))

# plot the sampling distribution
ggplot(data = sickness_sampling_dist, aes(x = prop_sick)) +
  geom_histogram(binwidth = 1e-3, color = "white", fill = "firebrick")

# find the standard error for the proportion of  
sickness_sampling_dist %>% 
  summarise(se = sd(prop_sick))
```


### Simulate testing   

Now let's test these people, and find the number that are    

1. healthy and diagnosed as healthy.       
2. healthy and diagnosed as sick.    
2. sick and diagnosed as healthy.   
3. sick and diagnosed as sick.   

Remember 

- That we use the [`case_when`](https://dplyr.tidyverse.org/reference/case_when.html) function to simulate conditional probabilities.    
- The test correctly reports someone uninfected as uninfected in 0.995 of tests.   
- The test correctly reports someone infected as infected in 0.94 of tests.    
<span style="color:lightgrey;">Note these answers will differ slightly from before because we are running the simulation again. </span>

```{r sim2, exercise=TRUE, eval = FALSE}
n_tests                    <- 10000
p_testsick_givensick       <- 0.94
p_testhealthy_givenhealthy <- 0.995
p_sick                     <- 0.02

sickness <- tibble(
  actual_status = sample(c("sick","healthy"), size = n_tests,  replace = TRUE, prob = c(p_sick, 1 - p_sick))) %>%
    group_by(actual_status) %>%
    mutate(test_status = case_when(
      actual_status == "sick" ~ sample(c("sick","healthy"),
            prob = c(___, 1 - ___), 
            replace = TRUE, size = n()),  
      actual_status == "healthy" ~  sample(c("sick","healthy"),
            prob = c(1 - ___, ___), 
            replace = TRUE,  size = n()))) %>%
  ungroup()

### Counting all combinations of outcomes
sickness %>%
  group_by(___, ___) %>%
  summarise(count = n(), .groups = "drop")



#### We will make a plot with the mosaic plot  function in base R. 
# These plots are nice, but hard to make in ggplot. 
# Build off of this code if you ever want to build one. 
# Mosaic plots are discussed in more detail in the Whitlock & Schluter textbook
table(sickness) %>% 
  mosaicplot(color = c("firebrick","blue"), 
             main = "", xlab = "actual status", ylab = "test status")

## Make your own better plot if you like!
```

```{r sim2-solution}
n_tests  <- 10000
p_testsick_givensick       <- 0.94
p_testhealthy_givenhealthy <- 0.995
p_sick                     <- 0.02

sickness <- tibble(
  actual_status = sample(c("sick","healthy"), size = n_tests,  replace = TRUE, prob = c(p_sick, 1 - p_sick))) %>%
    group_by(actual_status) %>%
    mutate(test_status = case_when(
      actual_status == "sick" ~  sample(c("sick","healthy"),
              prob = c(p_testsick_givensick , 1 - p_testsick_givensick ),
              replace = TRUE, size = n()),  
      actual_status == "healthy" ~ sample(c("sick","healthy"),
              prob = c(1 - p_testhealthy_givenhealthy, p_testhealthy_givenhealthy),
              replace = TRUE, size = n()))) %>%
  ungroup()

### Counting all combinations of outcomes
sickness %>%
  group_by(test_status, actual_status) %>%
  summarise(count = n(), .groups = "drop")

#### We will make a plot with the mosaic plot  function in base R. 
# These plots are nice, but hard to make in ggplot. 
# Build off of this code if you ever want to build one. 
# Mosaic plots are discussed in more detail in the Whitlock & Schluter textbook
table(sickness) %>% 
  mosaicplot(color = c("firebrick","blue"), 
             main = "", xlab = "actual status", ylab = "test status")

## Make your own better plot if you like!
```

### Processing simulation results to find Bayes' theorem  

Use the output of the simulations above to estimate the proportion of people that are diagnosed to be sick who are actually healthy. [just type in numbers below]

```{r calculator, exercise=TRUE, eval=FALSE}
___ / (___ + ___)
```

<div id="calculator-hint">
**Hint:** Divide the number of people who are diagnosed to be sick but are actually healthy by the number of people diagnosed as sick.
</div>


```{r bayes1, echo=FALSE}
question("In this example, the proportion of people that are diagnosed to be sick who are actually healthy is approximately",
  answer("0.005"),
  answer("0.04"),
  answer("0.20", correct = TRUE),
  answer("0.40"),
  answer("0.98"),
  allow_retry = TRUE
)
```


### Use Bayes' theorem

Use math to find the expected proportion of people who are diagnosed as sick who are actually healthy (assuming two percent of the population is sick)

```{r calculator2, exercise=TRUE, eval=FALSE}
p_testsick_givensick       <- 0.94
p_testhealthy_givenhealthy <- 0.995
p_sick                     <- 0.02

# use bayes theorem from last chapter 
# tmath here!!! 

```


### Which  better  helped you understand Bayes' theorem? 

The  simulation and counting or the math? Explain... 
