---
title: "Data In R Quiz"
author: "Yaniv Brandvain"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(ggmosaic)
iris <- tibble(iris)
ukgas<- data.frame(year = 1960:1986,matrix(c(UKgas), ncol = 4,byrow = TRUE))
names(ukgas) <- c("year","quarter_1", "quarter_2", "quarter_3", "quarter_4")
ukgas <- tibble(ukgas)
knitr::opts_chunk$set(echo = FALSE)
toad_data <- tribble(~individual,~species,~sound,~weight,~height,
        "a", "Bufo spinosus", "chirp", 2.0,	2,
        "b",	"bufo bufo",	"croak",	2.6,	3,
        "c",	"Bufo bufo",	"ribbit",	3.0,	2)
```

# Dealing with data Questions

###  Complete this quiz to evaluate your comprehension of the material
A similar, if not identical quiz should be completed for credit on canvas.

#### Q1: [`glimpse()`](https://tibble.tidyverse.org/reference/glimpse.html)

The `iris` dataset in `R` has floral measurements for fifty individuals each from three of `iris species`. Execute the code below, fix the mistake, and then answer the question which follows.

```{r glimpse, exercise=TRUE, eval = FALSE}
glimps(iris)
```


```{r glimpse-solution}
glimpse(iris)
```



```{r glimpseQ, echo=FALSE}
question("Species is of class character.",
    answer("TRUE"),
    answer("FALSE", correct = TRUE),
    allow_retry = TRUE,
    correct = random_praise(),
    incorrect = random_encouragement()
  )
```

  

#### Q2: [`filter()`](https://dplyr.tidyverse.org/reference/filter.html)     

Use the [`filter()`](https://dplyr.tidyverse.org/reference/filter.html) function to find out how many individuals in the dataset have a `Sepal.Length` greater than 7 units, and then answer the question.


```{r filter, exercise=TRUE, eval = FALSE}
filter(iris,  ___ > 7  )
```

```{r filter-solution}
filter(iris,  Sepal.Length > 7  )
## OR 
iris %>%
  filter(Sepal.Length > 7 )
```


```{r filterQ, echo=FALSE}
question_text("How many individuals Sepal.Length's greater than 7 units?",
    allow_retry = T,
    answer("12", correct = T),
    correct = random_praise(),
    incorrect = random_encouragement()
  )
```



#### Q3: Reproducibility 


```{r reproduce, echo=FALSE}
question("It is more important to save the code underlying an analysis than the result of the analysis.",
    answer("TRUE", correct = TRUE),
    answer("FALSE"),
    allow_retry = TRUE,
    correct = random_praise(),
    incorrect = random_encouragement()
  )
```

#### Q3-6: [`arrange()`](https://dplyr.tidyverse.org/reference/arrange.html)  again


**Q3:** Find the shortest character in the `starwars` data set [hint, you might want to [`glimpse()`](https://dplyr.tidyverse.org/reference/glimpse.html)  first.]

```{r arrange2, exercise=TRUE, eval = FALSE}
starwars_sortheight <- ____(starwars, ___) 
# Use the to just  show name and height of this new sorted data set
select(starwars_sortheight, ____, ___) 
```

```{r arrange2-solution}
starwars_sortheight <- arrange(starwars, height) 
# Use the to just  show name and height of this new sorted data set
select(starwars_sortheight, name, height) 
```


```{r arrange2Q, echo=FALSE}
question("Who is the shortest star wars character?",
    answer("Yoda", correct = TRUE),
    answer("R2-D2"),
    answer("Luke Skywalker"),
    answer("Yarael Poof"),
    allow_retry = TRUE,
    correct = random_praise(),
    incorrect = random_encouragement()
  )
```

**Q4:** 

```{r arrange2Qnochange, echo=FALSE}
question("After executing the code above, who is the first character in the `starwars` data set?",
    answer("Yoda"),
    answer("Luke Skywalker", correct = TRUE),
    allow_retry = TRUE,
    correct = random_praise(),
    incorrect = random_encouragement()
  )
```

**Q5:** 

```{r arrange2Qdoublecheck, echo=FALSE}
question("After executing the code above, who is the first character in the `starwars_sortheight` data set?",
    answer("Yoda", correct = TRUE),
    answer("Luke Skywalker"),
    allow_retry = TRUE,
    correct = random_praise(),
    incorrect = random_encouragement()
  )
```

**Q6:** 

```{r arrange2Qstillnochange, echo=FALSE}
question("After executing the code above, how many columns are in the ` `starwars_sortheight` dataset?",
    answer("87"),
    answer("14"),
    answer("2", correct = TRUE),
    allow_retry = TRUE,
    correct = random_praise(),
    incorrect = random_encouragement()
  )
```

