---
title: "Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(ggforce)
iris <- tibble(iris)
library(learnr)
library(datasauRus)
library(tidyverse)
library(ggmosaic)
library(cowplot)
iris <- tibble(iris)
knitr::opts_chunk$set(echo = FALSE)
library(ggthemes)
data_13 <- datasaurus_dozen %>%
  mutate(id = as.numeric(as.factor(dataset))) %>%
  select(-dataset) %>%
  mutate(id = letters[id])
```

# Review Data Summaries and Motivate Making Figures 


### The data

We are looking into `data_13` – a collection of dataset with three variables – `x`, `y`, and `id`, with thirteen different ids.

First, take a `glimpse()` of the `data_13` 

```{r glimpse, exercise=TRUE, eval = FALSE}
____(___)
```

```{r glimpse-solution, eval = FALSE}
glimpse(data_13)
```

### Summarize  

I will preview the skills you will learn in chapter 8 to have R calculate the mean and standard deviation for `x` and `y` by `id`. I am also giving you code to look into the correlation between these variables – a topic we’ll revisit in later chapters. For now, just click `run code`

```{r summaryz, exercise=TRUE, eval = FALSE}
data_13    %>%
  group_by(id)   %>%
  summarise(mean_x = mean(x),
       sd_x   = sd(x),
       mean_y = mean(y),
       sd_y   = sd(y),
       cor_xy = cor(x,y)) # a measure of the association between x and y
```





```{r message, echo=FALSE}
question("What is the take home message from these summaries?",
    answer("Across ids, summaries are similar for x, but different for y"),
    answer("Across ids summaries are different for x, but similar for y"),
        answer("All summaries are very similar across ids", correct = TRUE),
        answer("All summaries are very different across ids"),
    allow_retry = TRUE,
    correct = random_praise(),
    incorrect = random_encouragement()
  )
```

### Plot 

#### Jitterplot

Make a jittered plot of `data_13` by mapping `x` and `color` onto `id`, mapping `y` onto `y`, and specifying the appropriate geom for a jitterplot.

Leave all lines but the first and the geom on the second as is – they use advanced ggplot skills that we will introduce in the next chapter.



```{r jitterplot, exercise=TRUE, eval = FALSE}
ggplot(data_13, aes(x = id, __ = __ , __ = ___ ))      +
  geom____(width = .2, alpha = .5, show.legend = FALSE) +  # don't change these or worry about them, we'll revisit in the next chapter. 
  stat_summary(fun.data  = "mean_se", color = "black")        # don't change these or worry about them, we'll revisit later
```

```{r jitterplot-solution}
ggplot(data_13, aes(x = id, y = y ,color = id ))      +
  geom_jitter(width = .2, alpha = .5, show.legend = FALSE) +  # don't change these or worry about them, we'll revisit in the next chapter. 
  stat_summary(fun.data  = "mean_se", color = "black")        # don't change these or worry about them, we'll revisit later. 
```


```{r dinstinct, echo=FALSE}
question("Look at the jitterplot you just made. Which id has the most distinct shape? (that is which one looks leastlike the others)",
    answer("a"),
    answer("d"),
    answer("g", correct = TRUE),
    answer("h"),
    answer("i"),
    allow_retry = TRUE,
    incorrect = "Try again!")
```


#### Scatterplot


Use the skills you learned when we first introduced ggplot to present a scatterplot of the data, sectioned off by `id`.


```{r scatterplot, exercise=TRUE, eval = FALSE}
ggplot(___, aes(x = __, y = __)) +
  geom_point()                   + # leave blank
  facet_wrap(~__)
```

```{r scatterplot-solution}
ggplot(data_13, aes(x = x, y = y)) +
  geom_point()                   + # leave blank
  facet_wrap(~id)
```


```{r dino, echo=FALSE}
question("Which plot shows a dinosaur??",
    answer("a"),
    answer("d", correct = TRUE),
    answer("g"),
    answer("h"),
    answer("i"),
    allow_retry = TRUE,
    incorrect = "Try again!")
```
