---
title: "Probability I: Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(ggmosaic)
knitr::opts_chunk$set(echo = FALSE)
```

# Probabilistic thinking quiz  

## Background

Imagine there is some disease spreading, and people concerned about the disease take a test with the following behavior:   

- The test correctly reports someone uninfected as uninfected in 0.995 of tests.   
- The test correctly reports someone infected as infected in 0.94 of tests.  


```{r quiz1, echo=FALSE}
quiz(caption = "Concept questions",
    question("Imagine you do not have the disease, what is the probability that the test says you do?",
    answer("0.000"),
    answer("0.005", correct = TRUE),
    answer("0.060"),
    answer("0.94"),
    answer("0.995"),
    answer("1.000"),
    answer("Not enough info"),
    allow_retry = TRUE,
    incorrect = "Try again!"
  ),
  question("Which variables are independent? (check all that apply)",
    answer("Disease status is independent of test result"),
    answer("Being sick is independent of being not sick"),
    answer("None of the variables are independent of each other", correct = TRUE),
    answer("Not enough information"),
    allow_retry = TRUE,
    incorrect = "Try again!"
  ),
    question("Which variables are mutually exclusive? (check all that apply)",
    answer("Being sick and having a test say you are not sick are mutually exclusive"),
    answer("Being sick and not being sick are mutually exclusive",  correct = TRUE),
    answer("Being healthy and having a test say you are sick are mutually exclusive"),
    answer("Having the test say you are sick and having the test say you are not sick are mutually exclusive",  correct = TRUE),
    answer("None of the variables are mutually exclusive"),
    answer("Not enough information"),
    allow_retry = TRUE,
    incorrect = "Try again!"
  ),
  question("Which is the explanatory variable and which is the response?",
    answer("Sick is Explanatory, Test is response",  correct = TRUE),
    answer("Test is Explanatory, Sick is response"),
    answer("Not enough info"),
    allow_retry = TRUE,
    incorrect = "Try again!"
  ),
    question("How MIGHT results of this test paint a biased view of the proportion of the population that has the disease? (check all that apply)",
    answer("Sick people are more likely to be tested than healthy people (contributing to a case positivity rate greater than the proportion of the population that is sick)",  correct = TRUE),
    answer("Risk takers recklessly exposing themselves to the disease might not care to know if they are infected. (contributing to a case positivity rate lower than the proportion of the population that is sick)",  correct = TRUE),
    answer("False positives could upwardly bias our estimate of the proportion sick", correct = TRUE),
    answer("There is no opportunity for bias, the test is highly accurate and the people conducting the tests are morally upright medical professionals."),
    answer("Not enough information"),
    allow_retry = TRUE,
    incorrect = "Try again!"
  )
)
```


**Consider the question, above:**    
- Can you think of any biases not listed above?   
- Which bias do you think is responsible for the greatest difference between disease rate and test positivity rate? Why?   


### Maths 

Above we imagined there is some disease spreading, and people concerned about the disease take a test with the following behavior:   

- The test correctly reports someone uninfected as uninfected in 0.995 of tests.   
- The test correctly reports someone infected as infected in 0.94 of tests.  

Now assume that **two in one hundred people have the disease.** 




#### Right diagnosis   


**Right diagnosis** Imagine that in one day, 10 healthy people get tested. Find the probability that  all ten get a (correct) negative result (to four digits past the decimal)).   


```{r calculator0, exercise=TRUE, eval=FALSE}
p_testsick_givensick       <- 0.94
p_testhealthy_givenhealthy <- 0.995
p_sick                      <- 0.02
```




<div id="calculator0-hint">
**Hint:** We get correctly diagnosed as healthy by getting one true positive AND another true positive, and another... 

Because we multiply each independent probability, the probability of $n$ consecutive outcomes, each with probability $p$ is  $p^n$.

More generally, the probability of a sequence of outcomes is $\prod_{i=1}^{n}  p_i$, with   $p^n$ as a special case for when the probability of each event in your sequence is the same.  
</div>




```{r exponent, echo = FALSE}
question_text("All ten get a (correct) negative result (to four digits past the decimal))",
 answer("0.9511", correct = TRUE),
 answer( ".9511", correct = TRUE),
    allow_retry = TRUE,
    correct = random_praise(),
    incorrect = random_encouragement()
)
```


#### Wrong diagnosis   

**Wrong diagnosis** Imagine that in one day, 10 healthy people get tested. Find the probability that   at lease one  gets an (incorrect) positive result (to four digits past the decimal)   


For math   

```{r calculator1, exercise=TRUE, eval=FALSE}
p_testsick_givensick       <- 0.94
p_testhealthy_givenhealthy <- 0.995
p_sick                      <- 0.02
```


<div id="calculator1-hint">
**Hint:** The probability of not something equals 1 minus the probability of that thing. 

So this should be 1 - the answer above.  
</div>




```{r not, echo = FALSE}
question_text("At lease one  gets an (incorrect) positive result (to four digits past the decimal)",
 answer("0.0489", correct = TRUE),
 answer( ".0489", correct = TRUE),
    allow_retry = TRUE,
    correct = random_praise(),
    incorrect = random_encouragement()
)
```

### Use the mutliplication rule to find 

The proportion of people who are  healthy but diagnosed with the disease  (to four digits past the decimal)"

```{r calculator2, exercise=TRUE, eval=FALSE}
p_testsick_givensick       <- 0.94
p_testhealthy_givenhealthy <- 0.995
p_sick                      <- 0.02
```


<div id="calculator2-hint">
**Hint:** Multiply the probability that someone is healthy by the probability a healthy person gets a false positive.    
</div>



```{r multiplication, echo = FALSE}
question_text("The proportion of people who are  healthy but diagnosed with the disease  (to four digits past the decimal)",
 answer("0.0049", correct = TRUE),
 answer( ".0049", correct = TRUE),
    allow_retry = TRUE,
    correct = random_praise(),
    incorrect = random_encouragement()
)
```

### Use the law of total probability to find 

Use math to find the proportion of people diagnosed with the disease (to four digits)

```{r calculator3, exercise=TRUE, eval=FALSE}
p_testsick_givensick       <- 0.94
p_testhealthy_givenhealthy <- 0.995
p_sick                      <- 0.02
```


<div id="calculator3-hint">
**Hint:** People can be diagnosed if they are healthy and get a FALSE positive OR if they are sick and get a true positive. Theses are mutually exclusive so we add the probabilities without fear of double counting.
</div>



```{r totalProb, echo = FALSE}
question_text("The proportion of people diagnosed with the disease (to four digits)",
 answer("0.0237", correct = TRUE),
 answer( ".0237", correct = TRUE),
  allow_retry = TRUE,
  correct = random_praise(),
  incorrect = random_encouragement()
)
```


### Use Baye's theorem  to find 

The proportion of people diagnosed with the disease who actually have it (with four digits past the decimal). That is: number diagnosed and sick / number diagnosed

```{r calculator4, exercise=TRUE, eval=FALSE}
p_testsick_givensick       <- 0.94
p_testhealthy_givenhealthy <- 0.995
p_sick                      <- 0.02
```





```{r bayes, echo = FALSE}
question_text("The proportion of people diagnosed with the disease who actually have it (with four digits past the decimal). That is: number diagnosed and sick / number diagnosed ",
 answer("0.7932", correct = TRUE),
 answer( ".7932", correct = TRUE),
  allow_retry = TRUE,   
  correct = random_praise(),
  incorrect = random_encouragement()
)
```



